Lactate transport inhibition therapeutically reprograms fibroblast

### 2024-02-16 manuscript/manuscript.qmd
working on isotope labeling data

>>>>

## MCT inhibition reprograms lung metabolism *in vivo*

To investigate metabolic perturbations following lactate transporter inhibition *in vivo*, we next performed metabolomic profiling of lung and plasma from mice treated with bleomycin and MCT inhibitors (**Supplemental Figure 9.1**). While bleomycin treatment substantially altered the lung metabolic profile, treatment with AZD3965 or VB124 had little additional impact on total metabolite levels. Unlike bleomycin alone or following AZD3965 treatment, VB124 treatment was associated with significant changes in several circulating metabolites (**Supplemental Figure 9.1I**). Among these, VB124 significantly increased circulating lactate, while AZD3965 was associated with decreased circulating lactate (**Figure 9A**). This pattern was also reflected in the lung, where only modest changes were observed following bleomycin, contrary to prior reports [@kottmann2012]. Given the impact of our treatments on circulating lactate, we calculated the ratio of lung-to-plasma lactate. This analysis demonstrated the expected increase in lactate in fibrotic lungs with levels returning to baseline following the addition of VB124 (**Figure 9A**).

We were surprised to observe a relative paucity of metabolic changes in whole lung lysates following MCT inhibition compared to our *in vitro* studies. We hypothesized that this might be a consequence of the cellular heterogeneity of MCT expression in the fibrotic lung. To examine this hypothesis, we performed multi-isotope imaging mass spectrometry (MIMS) of mouse lungs following treatment with ^15^N-proline and ^2^H-glucose. MIMS enables the quantification of stable isotope tracer uptake into tissue biomass with subcellular spatial resolution [@steinhauser2012; @narendra2020]. For three days prior to tissue collection, mice received twice daily intraperitoneal injections of 5 mg ^15^N-proline as a tracer for tissue fibrosis and 50 mg ^2^H-glucose as a metabolic tracer. Lung tissue sections were subsequently imaged by nanoscale seconary ion mass spectrometry to quantify spatially resolved isotope tracer uptake (**Figure 9B**).
